---
title: "Vanderbuilt Excel Reproduction - Notebook"
author: "Martin Frigaard"
date: "2018-04-12"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
---


```{r knitr_init, echo=FALSE, cache=FALSE}
require(knitr)
require(rmdformats)
require(magrittr)
require(tidyverse)
## Global options
options(max.print = "75")
opts_chunk$set(size = "small",
               prompt = FALSE,
               tidy = FALSE,
               message = FALSE,
               warning = FALSE)
opts_knit$set(width = 75)
# file time
file_time <- lubridate::now(tzone = "US/Pacific-New") %>% 
    as.character.Date() 
# what is the file title? ----
file_title <- c("vand excel repro")
# what is the file version? ----
file_version <- c("1.1") # file version 1.1
```

**PACKAGES:**  

```{r packages, message=FALSE, warning=FALSE}
library(dplyr) # Data wrangling, glimpse(50) and tbl_df().
library(ggplot2) # Visualise data.
library(lubridate) # Dates and time.
library(readr) # Efficient reading of CSV data.
library(stringr) # String operations.
library(tibble) # Convert row names into a column.
library(tidyr) # Prepare a tidy dataset, gather().
library(magrittr) # Pipes %>%, %T>% and equals(), extract().
library(tidyverse) # all tidyverse packages
library(mosaic) # favstats and other summary functions
library(fs) # file management functions
library(shiny) # apps 
```

**File name:**

```{r create_file_name, echo=FALSE}
file_title <- stringr::str_to_lower(
    stringr::str_replace_all(file_title, "\\W", "_"))
file_name <- paste0(file_version,"-",file_title, ".Rmd")
file_name
```

**HEADER:**  


* **File name:** `r file_name`
* **Created date:** `r Sys.Date()`    
* **R version:** `r R.version.string`  

# Motivation

This notebook covers the reproduction of the analyses sent to Vanderbilt in the following document:

## Load custom function

This will come in handy for summarizing data 

```{r num_vars_summ}
# num_vars_summ function ----------------------------------------------
# this summarizes a numerical variable (n, na(missing, mean, median, sd,
# variance, min, max, and se) when given a data frame or tibble and
# variable: q14_rpt_fun(df, var)
num_vars_summ <- function(df, expr) {
  expr <- enquo(expr) # turns expr into a quosure
  summarise(df,
          n = sum((!is.na(!!expr))), # non-missing
          na = sum((is.na(!!expr))), # missing
          mean = mean(!!expr, na.rm = TRUE), # unquotes mean()
          median = median(!!expr, na.rm = TRUE), # unquotes median()
          sd = sd(!!expr, na.rm = TRUE), # unquotes sd()
          variance = var(!!expr, na.rm = TRUE), # unquotes var()
          min = min(!!expr, na.rm = TRUE), # unquotes min()
          max = max(!!expr, na.rm = TRUE), # unquotes max()
          se = sd/sqrt(n)) # standard error
}
```


`Excel/VanderbiltUpdated.xlsx`

```{r VanderbiltUpdated.xlsx}
fs::dir_ls("Excel")
```

This file has four tabs: SEG Output, Surveillance Study Criteria, MARD, ModBA Plot

## SEG Output

This tab contains the image from the previous script that re-creates the heat map, and the data used to re-create this image. 

### Data for SEG Output

This will load the data from the SEG output

```{r load_data}
# fs::dir_ls("Data/Raw") %>% writeLines()
RiskPairData <- read_csv("Data/Raw/RiskPairData.csv")
LookUpRiskCat <- read_csv("Data/Raw/LookUpRiskCat.csv")
SampleData <- read_csv("Data/Raw/SampleData.csv")
```

```{r RiskPairData}
RiskPairData %>% glimpse(60)
```

```{r SampleData}
SampleData %>% glimpse(60)
```

```{r LookUpRiskCat}
LookUpRiskCat %>% glimpse(60)
```

## Rename `BMP_result` to `REF` and `POC_Result` to `BGM`

This will change the current `BMP_Result` (which stands for Basic Metabolic Panel) to `REF`.

```{r wrangle}
# rename POC_Result, BMP_Result
# SampleData <- SampleData %>% dplyr::rename(
#     BGM = POC_Result,
#     REF = BMP_Result)

# Pair type variables ----
SampleData <- SampleData %>% 
                dplyr::mutate(
                    pairtype_gt600 = dplyr::case_when(
                                    REF > 600 ~ "BMP > 600",
                                    TRUE ~ NA_character_),
                    pairtype_lt21 = dplyr::case_when(
                                    REF < 21 ~ "BMP < 21",
                                    TRUE ~ NA_character_),
                    pair_type = dplyr::case_when(
                                    BGM < REF ~ "POC < BMP",
                                    BGM == REF ~ "POC = BMP",
                                    BGM > REF ~ "POC > BMP",
                                    TRUE ~ NA_character_))




# These counts are below (I use the original data set `SampleData`)
# this is a static output for pair type--include knitr
## 1.) Pair Type Table == == == ==  ----
PairType <- SampleData %>% 
    dplyr::count(pair_type)
## 2.) Pair Type Greater than 600 == == == ==  ----
PairTypeGT600 <- SampleData %>% 
    dplyr::count(pairtype_gt600) %>% 
    dplyr::rename(`REF greater than 600` = pairtype_gt600)
## 3.) Pair Type Less than 21 == == == ==  ----
PairTypeLT21 <- SampleData %>% 
    dplyr::count(pairtype_lt21) %>% 
    dplyr::rename(`REF less than 21` = pairtype_lt21)
```

## Create risk category table 

```{r SEG_pair_type}
SampleData <- SampleData %>% 
dplyr::mutate(
        measured = poc_result,
        reference = ref_result,
        seg_pair_type = 
        dplyr::case_when( 
            measured > 600 ~ "out of range",
            reference > 600 ~ "out of range",
            measured <= 20  ~ "out of range",
            reference <= 20 ~ "out of range",
            reference > measured ~ "BGM below ref",
            reference < measured ~ "BGM above ref",
            measured == reference ~ "BGM equal to ref"))

```

```{r RiskPairData_glimpse}
RiskPairData %>% glimpse(60)
```

Now join the `RiskPairData` data to our `SampleData`.

```{r SampleData_join_RiskPairData}
SampleDataRiskPairData <- dplyr::inner_join(x = SampleData, 
                   y = RiskPairData, 
                   by = c("measured", "reference"))
SampleDataRiskPairData %>% glimpse(75)
```

## Create `risk_cat` vector

This assigns the appropriate risk category to the risk factor value. 

```{r abslb_vec}
abslb_vec <- LookUpRiskCat$abslb
abslb_vec
```

This comes in handy for the lookup table (with `base::findInterval()` function), although not necessary.

Now we will use the `findInterval()` function to create a new variable based on the `abslb_vec` levels. 

```{r create_risk_cat} 
# test ----
SampleDataRiskPairData %>% 
    dplyr::mutate(risk_cat = base::findInterval(x = abs_risk, # the abs_risk absolute value
                                   vec = abslb_vec, # the vector we created
                                   left.open = TRUE) - 1) %>% dplyr::count(risk_cat) 
# assign ----
SampleDataRiskPairData <- SampleDataRiskPairData %>% 
    dplyr::mutate(risk_cat = base::findInterval(x = abs_risk, # the abs_risk absolute value
                                   vec = abslb_vec, # the vector we created
                                   left.open = TRUE) - 1) 
SampleDataRiskPairData %>% 
    group_by(risk_cat) %>% 
    num_vars_summ(abs_risk)
```

Now give these variables a text value in `risk_cat_txt`

```{r risk_cat_txt}
# test -----
SampleDataRiskPairData %>% 
    dplyr::mutate(risk_cat_txt = 
                      dplyr::case_when(abs_risk < 0.5 ~ "None",
                                       abs_risk >= 0.5 & abs_risk <= 1 ~ "Slight, Lower", 
                                       abs_risk > 1 & abs_risk <= 1.5 ~ "Slight, Higher", 
                                       abs_risk > 1.5 & abs_risk <= 2.0 ~ "Moderate, Lower",
                                       abs_risk > 2 & abs_risk <= 2.5 ~ "Moderate, Higher",
                                       abs_risk > 2.5 & abs_risk <= 3.0 ~ "Severe, Lower",
                                       abs_risk > 3.0 & abs_risk <= 3.5 ~ "Severe, Higher", 
                                       abs_risk > 3.5 ~ "Extreme")) %>% 
                    dplyr::count(risk_cat_txt)
# assign ----
SampleDataRiskPairData <- SampleDataRiskPairData %>% 
    dplyr::mutate(risk_cat_txt = 
                      dplyr::case_when(abs_risk < 0.5 ~ "None",
                                       abs_risk >= 0.5 & abs_risk <= 1 ~ "Slight, Lower", 
                                       abs_risk > 1 & abs_risk <= 1.5 ~ "Slight, Higher", 
                                       abs_risk > 1.5 & abs_risk <= 2.0 ~ "Moderate, Lower",
                                       abs_risk > 2 & abs_risk <= 2.5 ~ "Moderate, Higher",
                                       abs_risk > 2.5 & abs_risk <= 3.0 ~ "Severe, Lower",
                                       abs_risk > 3.0 & abs_risk <= 3.5 ~ "Severe, Higher", 
                                       abs_risk > 3.5 ~ "Extreme"))
```

## Surveillance Study Criteria

This sheet has the study criteria and breaks the table categories into *ISO Range* values. 

These are defined as, 

> "Difference between POC and BMP as % of BMP for BMP >100 mg/dL and in mg/dL for BMP ≤ 100 mg/dL"

Assuming each category getting its percentage from the total `7857`, we need to calculate this number from the `SampleData` data frame. 

```{r}
SampleData %>% nrow()
```

I will create a `iso_diff` column between POC and BMP,

```{r ISO_diff}
# SampleData$POC_Result %>% class()
# SampleData$BMP_Result %>% class()
# format the BMP_Result and POC_result
SampleData$ref_result <- as.double(SampleData$ref_result)
SampleData$poc_result <- as.double(SampleData$poc_result)
SampleData <- SampleData %>% 
    dplyr::mutate(iso_diff = if_else(ref_result >= 100, # condition 1
                                    100*abs(poc_result - ref_result)/ref_result, # TRUE 1
                                        if_else(ref_result < 100, # 
                                                abs(poc_result - ref_result),
                                                        NA_real_),
                                                            NA_real_)) # FALSE
SampleData %>% 
    num_vars_summ(iso_diff)
```

> What is the number of observations with an ISO range less than 5, 10, 15?

If `ref` is greater than 100, then calculate % of `ref` 

```{r iso_range}
SampleData %>%
    dplyr::mutate(
        iso_range =
            dplyr::case_when(iso_diff <= 5 ~ "<= 5% or 5 mg/dL",
                      iso_diff > 5 & iso_diff <= 10 ~ "> 5 - 10% or mg/dL",
                      iso_diff > 10 & iso_diff <= 15 ~ "> 10 - 15% or mg/dL",
                      iso_diff > 15 & iso_diff ~ "> 15% or 15 mg/dL")) %>% count(iso_range)
SampleData <- SampleData %>%
    dplyr::mutate(
        iso_range =
            dplyr::case_when(iso_diff <= 5 ~ "<= 5% or 5 mg/dL",
                      iso_diff > 5 & iso_diff <= 10 ~ "> 5 - 10% or mg/dL",
                      iso_diff > 10 & iso_diff <= 15 ~ "> 10 - 15% or mg/dL",
                      iso_diff > 15 & iso_diff ~ "> 15% or 15 mg/dL")) 
```


## The "Inverse of the Cumulative Binomial Distribution" (or `BINOM.INV`)

The `BINOM.INV` function in Excel is used to calculate the "the smallest value for which the cumulative binomial distribution is greater than or equal to a criterion value."

Failures will be in the `>=15% or 15 mg/dL` category of `ISO_range`--subtract from total to get success. 

```{r successes_trials}
successes <- nrow(SampleData) - base::nrow(dplyr::filter(SampleData, iso_range == "> 15% or 15 mg/dL"))
successes
trials <- nrow(SampleData)
trials
# Trials could also be the number of rows in the total data frame 
# (with out of range values)
```

Then a `prob` argument is applied, or the probability of a success on each trial. 

```{r prob}
prob <- 0.95
```

Finally an Alpha or `criterion` value

```{r criterion}
criterion <- 0.05
```

```{r binom.test}
binom.test(x = successes, 
           n = trials, 
           p = prob, 
           conf.level = 0.95)
```

## MARD

Graph two quantities: 

`POC - REF / REF` -> call this `rel_diff`

`|rel_diff|` -> call this `abs_rel_diff`


```{r create_rel_diff_Summary}
SampleData <- SampleData %>% 
    mutate(rel_diff = (poc_result - ref_result)/ref_result,
           abs_rel_diff = abs(rel_diff))
Summary <- tribble(
             ~`Total`,                 ~`Bias`,                     ~`MARD`,                  ~`CV`, 
       nrow(SampleData), mean(SampleData$rel_diff), mean(SampleData$abs_rel_diff),  sd(SampleData$rel_diff))
knitr::kable(Summary)
```

> adjust digits!

Lower and upper bound from excel!

Bias-1.96*sd
Bias+1.96*sd

```{r add_bias_to_Summary}
Summary <- Summary %>% 
    add_column(`Lower Limit` = Summary$Bias - 1.96*Summary$CV)
Summary <- Summary %>% 
    add_column(`Upper Limit` = Summary$Bias + 1.96*Summary$CV)
Summary
```

## Add the squared relative difference (`sq_rel_diff`) to data frame

```{r sq_rel_diff}
SampleData <- SampleData %>% 
    mutate(sq_rel_diff = rel_diff^2)
```

## Bland Altman Plot (modified)

This plot needs the scales adjusted, and the Bland-Altman ranges set up from the tibble below. 

```{r}
SampleData %>% 
    ggplot(aes(x = ref_result,
               y = rel_diff)) +
    geom_point()
```

### Create the `BlandALtRef`

Create the modified Bland-altman plot range reference values. 

```{r BlandALtRef}
BlandALtRef <- tribble(
~`Ref`,	~`UB`,	~`LB`,
20.00,	0.750,	-0.750,
30.00,  0.500,	-0.500,
40.00,  0.375,	-0.375,
50.00,	0.300,	-0.300,
60.00,	0.250,	-0.250,
70.00,	0.214,	-0.214,
90.00,	0.167,	-0.167,
100.00,	0.150,	-0.150,
400.00,	0.150,	-0.150,
450.00,	0.150,	-0.150,
500.00,	0.150,	-0.150
)
BlandALtRef
```

```{r}
library(blandr)
# Generates two random measurements
measurement1 <- rnorm(100)
measurement2 <- rnorm(100)

# Generates a ggplot
# Do note the ggplot function wasn't meant to be used on it's own
# and is generally called via the bland.altman.display.and.draw function


# Passes data to the blandr.statistics function to generate Bland-Altman statistics
statistics.results <- blandr.statistics(measurement1, measurement2)

# Generates a ggplot, with no optional arguments
blandr::blandr.plot.ggplot(statistics.results)

# Generates a ggplot, with title changed
blandr.ggplot(statistics.results , plotTitle = "Bland-Altman example plot" )

# Generates a ggplot, with title changed, and confidence intervals off
blandr.ggplot(statistics.results , plotTitle = "Bland-Altman example plot" ,
ciDisplay = FALSE , ciShading = FALSE )
```




