---
title: "SEG Shiny Heatmap - binomialTable() function Error (version 1.4)"
author: "Martin Frigaard"
date: "`r Sys.Date()`" 
output:
  rmdformats::readthedown:
    highlight: kate
---


```{r setup, include=FALSE}
require(tidyverse)
require(mosaic)
require(magrittr)
library(knitr)
library(rmdformats)
## Global options
options(max.print = "75")
knitr::opts_chunk$set(
  size = "small",
  cache = FALSE,
  prompt = FALSE,
  tidy = FALSE,
  comment = NA,
  message = TRUE,
  warning = TRUE
)
opts_knit$set(width = 75)
library(dplyr) # Data wrangling, glimpse(50) and tbl_df().
library(ggplot2) # Visualise data.
library(lubridate) # Dates and time.
library(readr) # Efficient reading of CSV data.
library(stringr) # String operations.
library(tibble) # Convert row names into a column.
library(tidyr) # Prepare a tidy dataset, gather().
library(magrittr) # Pipes %>%, %T>% and equals(), extract().
library(tidyverse) # all tidyverse packages
library(mosaic) # favstats and other summary functions
library(fs) # file management functions
library(shiny) # apps
library(datapasta) # for pasting tibbles
```

**HEADER:**  

* **Created date:** `r Sys.Date()`    
* **R version:** `r R.version.string`  

***

## Load test data 

To start with, we will upload the complete data we will use to make sure the wrangling/calculations are correct. The first step is to name the entire `VanderbiltComplete.csv` data set `SampMeasData`.


```{r SampMeasData_vanderbilt_repo}
# 1.6.1 load the vanderbilt data set ---- ---- ---- ----
# writeLines(fs::dir_ls(app_data_path))
vanderbilt_repo <- "mjfrigaard/SEG_shiny/master/Data/VanderbiltComplete.csv"
SampMeasData <- read_csv(paste0(github_root, vanderbilt_repo))
SampMeasData %>% dataShape()
# Observations: 9891
# Variables: 2
```

Load the `segTable()` function from version 1.4.

```{r segTable_v1.4.R}
# writeLines(fs::dir_ls("."))
source("segTable_v1.4.R")
```

This was the previous `binomialTable()` function--version 1.3.

```{r binomialTable, eval=TRUE}
# 4.0 binomialTable ---- ----  ---- ----  ---- ----  ---- ----
binomialTable <- function(dataset) {
  compliant_pairs <- nrow(dataset) - base::nrow(filter(
    dataset,
    iso_range %in% "> 15% or 15 mg/dL"
  )) %>%
    tibble::as_tibble() %>%
    dplyr::rename(`Compliant Pairs` = value)

  CompliantPairs <- compliant_pairs %>%
    dplyr::mutate(
      `Compliant Pairs %` = 100 * `Compliant Pairs` / nrow(dataset),
      `Compliant Pairs %` = base::round(`Compliant Pairs %`, digits = 2),
      `Compliant Pairs %` = base::paste0(`Compliant Pairs %`, "%")
    )

  # 4.0.1 create probability ---- ----  ---- ----
  prb <- 0.95
  p_value <- 0.05
  df_size <- nrow(dataset)

  # 4.0.2 create criterion ---- ---- ---- ----
  qbinom_table <- broom::tidy(qbinom(
    p = p_value,
    size = df_size,
    prob = prb,
    lower.tail = TRUE,
    log.p = FALSE
  )) %>%
    # clean up this variable in the tibble for display
    dplyr::rename(`Lower Bound for Acceptance` = x)

  QbinomTable <- qbinom_table %>%
    dplyr::mutate(
      `Lower Bound for Acceptance %` = 100 * `Lower Bound for Acceptance` / nrow(dataset),
      `Lower Bound for Acceptance %` = base::round(`Lower Bound for Acceptance %`, digits = 2),
      `Lower Bound for Acceptance %` = base::paste0(`Lower Bound for Acceptance %`, "%")
    )

  BinomialTest6 <- bind_cols(CompliantPairs, QbinomTable)
  # 4.0.3 add the test for acceptance text ---- ----  ---- ----
  BinomialTest6 %>% dplyr::mutate(
    Result =
      if_else(condition = `Compliant Pairs` < `Lower Bound for Acceptance`,
        true = paste0(
          BinomialTest6$`Compliant Pairs`[1],
          " < ",
          BinomialTest6$`Lower Bound for Acceptance`[1],
          " - Does not meet BGM Surveillance Study Accuracy Standard"
        ),
        false = paste0(
          BinomialTest6$`Compliant Pairs`[1],
          " < ",
          BinomialTest6$`Lower Bound for Acceptance`[1],
          " - Meets BGM Surveillance Study Accuracy Standard"
        )
      )
  )
}
```

Test it. 

```{r 4.0.4_test_binomialTable}
# there is a data file in the Data folder that this function will clean
binomialTable(SampMeasData)
```

### 4.0.5 - Error message `Warning: 'tidy.numeric' is deprecated.`

The following error came up in `broom`

```{r 4.0.6_error_tidy.numeric, error=TRUE}
# 4.1 create probability
prb <- 0.95
p_value <- 0.05
df_size <- nrow(SampMeasData)
qbinom_table <- broom::tidy(qbinom(
  p = p_value,
  size = df_size,
  prob = prb
)) %>%
  # clean up this variable in the tibble for display
  dplyr::rename(`Lower Bound for Acceptance` = x)
qbinom_table
```

Replace this with `tibble::as_tibble()`

```{r 4.0.7_fix_error_broom_tidy}
# 4.1 create probability
prb <- 0.95
p_value <- 0.05
df_size <- nrow(SampMeasData)
qbinom_table <- qbinom(
  p = p_value,
  size = df_size,
  prob = prb
) %>%
  tibble::as_tibble() %>%
  # clean up this variable in the tibble for display
  dplyr::rename(`Lower Bound for Acceptance` = value)
qbinom_table
```


This fixed and was updated in version 1.4.
